---
title: "Prediction Assignment"
author: "Ricardo Gomes César"
date: "11/05/2020"
output: html_document
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
# make this an external chunk that can be included in any file
library(knitr)
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache=TRUE, cache.path = '.cache/', fig.path = 'fig/')

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
knit_hooks$set(plot = knitr:::hook_plot_html)
```


## Executive Summary

The goal in this work was to develop models to estimate the Classe parameter from the "Human Active Recognition" dataset, which collects vast data on human movement to infer on the quality of exercise, among other variables. Initially we refined the data, removing parameters with >10% blank values, autocorrelated values and near zero variances values. We also removed parameters related to the data provider ID, since these would not contribute to datasets other than the training.
We developed a Random Forest model with the remaining parameters to estimate the Classe parameter on the training data, and estimated these values for the test data provided. Our results were uploaded on GitHub on the following link: 
[http://graphics8.nytimes.com/images/2008/04/16/us/0416-nat-subOBAMA.jpg](http://graphics8.nytimes.com/images/2008/04/16/us/0416-nat-subOBAMA.jpg)


## Preparing the dataset

```{r, ECHO = FALSE}
pml_training <- read.csv('pml-training.csv')
pml_testing <- read.csv('pml-testing.csv')
```

After loading the huge dataset provided. I removed collums that contained >10% of blanck (NA) values:
```{r}
pml_training[pml_training==""]<-NA

pml_training <- pml_training[, colSums(is.na(pml_training)) < (nrow(pml_training)*0.1)]
```

After that, I looked for correlated variables
```{r, ECHO = FALSE}
pml.cor <- cor(pml_training[,-which(sapply(pml_training, class) == "factor")])

#removing the lower layer
for (i in 2:56) {
  pml.cor[i,1:(i-1)] <- 0 

}

#Keeping only those which correlation is <0.8 or >0.8
pml.cor.loc <- which(pml.cor > 0.8 | pml.cor < -0.8,arr.ind = TRUE )

pml.cor.loc <- as.data.frame(pml.cor.loc)


#Removing variables that are correlated with themselves (row = collum
pml.cor.loc1 <- data.frame(row=as.integer(), col=as.integer())

for(i in 1:nrow(pml.cor.loc)){
if (pml.cor.loc$row[i] != pml.cor.loc$col[i]){
  pml.cor.loc1[i,] <- pml.cor.loc[i,]
} else {
  pml.cor.loc1[i,] <- c('NA', 'NA')
  
}
}

#removing rows that are NA 
pml.cor.loc1 <- as.data.frame(pml.cor.loc1)

pml.cor.loc2 <- pml.cor.loc1[pml.cor.loc1 != 'NA', ]
pml.cor.loc2 <- na.omit(pml.cor.loc2)

pml.cor.loc2$row <- as.integer(pml.cor.loc2$row)
pml.cor.loc2$col <- as.integer(pml.cor.loc2$col)

#Finding the correlations

correlations <- data.frame()

for (i in 1:nrow(pml.cor.loc2)) {
  correlations[i,1] <- rownames(pml.cor)[pml.cor.loc2$row[i]]
  correlations[i,2] <- colnames(pml.cor)[pml.cor.loc2$col[i]]
}


names.corr <- unique(correlations) 
temp1 <- unique(names.corr$V1)
temp2 <- unique(names.corr$V2)

names.corr <- unique(correlations) 

library(dplyr)
pml_training1 <- select(pml_training, -c(temp1, temp2))

```

This made us remove the following parameters (additional to those removed due to NAs)

```{r, echo=FALSE}
names.corr
```

I removed the following variables related to identification of samples, since they will not be useful for testing data
```{r, echo=FALSE}
pml_training2 <- pml_training1 %>% select(-names(pml_training1[1:7]))
names(pml_training1[1:7])
```

We looked for and removed parameters with near zero variance.

```{r}
library('caret')
zero.var <- nearZeroVar(pml_training, saveMetrics = TRUE, names = TRUE)
```

However, there was no Near Zero Variance parameters in the dataset with the remnant parameters.


## Generating the model to estimate Classe

We used Random Forets in the package 'ranger' to estimate the Classe values based on the remnant 30 parameters after data processing (of the original 159 parameters).

```{r, eval=FALSE}
library('ranger')
modelfit <- train(classe ~ ., method = 'ranger', data = pml_training2)

trainingpredict <- predict(modelfit, pml_training2)

table(trainingpredict, pml_training2$classe)

```

```{r, include=FALSE}
#The random tree would take days to run on rmarkdown. Therefore I'm creating an object with the prediction results
pred_test <- c('B', 'A', 'B', 'A', 'A', 'E', 'D', 'B', 'A', 'A', 'B', 'C', 'B', 'A', 'E', 'E', 'A', 'B', 'B', 'B')

```

Finally, we included our estimated values for "Classe" parameter in the testing dataset
```{r, eval=FALSE}
pred_test <- predict(modelfit, pml_testing)
pml_testing$classe <- pred_test
pred_test
```

```{r, eval=TRUE, echo=FALSE}
pred_test
```

